Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2024-12-08 latex/intro.tex
Submitted minor revision

As another example illustrating the differences between replication
and clusterization, consider an experiment involving a chemical
reaction at different temperatures. An investigator may choose to
\emph{replicate} the experiment at $10^\circ C$ and $30^\circ
C$. Within a single replication, the investigator may also choose to
take measurements at different times. If the investigator chooses to
take two measurements at the same time (within a single replication),
then the investigator has utilized a \emph{clustered} design. While
replication is intuitively sensible for this chemical experiment,
clusterization is not.

@@

Researchers of inverse problems widely agree that measurement
clusterization is undesirable \cite{fedorov1996, nyberg2012,
  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
exploration of various remedies to address this issue. One approach
involves merging close measurements \cite{fedorov1997}; however, this
strategy merely overlooks the phenomenon of measurement
clusterization. An alternative solution lies in
\emph{clusterization-free design}s, where measurement locations are
deliberately chosen to be distant from one another. This can be
achieved by imposing distance constraints between measurements or by
introducing correlated errors that account for both observation error
and model misspecification \cite{Ucinski05}. For instance, in the
context of time-series analysis for pharmacokinetic experiments,
measurement clusterization can be mitigated by incorporating the
modeling of auto-correlation time within the noise terms
\cite{nyberg2012}.

>>>>

As another example illustrating the differences between replication
and clusterization, consider an experiment involving a chemical
reaction conducted at different temperatures. An investigator may
choose to repeat the experiment at $10^\circ C$ and $30^\circ C$,
giving rise to \emph{replication}. Within a single experiment, the
investigator may also sample the concentrations of the reagents
through the course of the experiment. Simultaneous sampling (within a
single experiment) gives rise to \emph{clusterization}. While
replication is intuitively sensible for this chemical experiment,
clusterization is not.

@@

Measurement clusterization is generally considered undesirable
\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
  neitzel2019sparse}. It is counterintuitive for a so-called "optimal"
design to prioritize marginally improving accuracy in a small region
over taking measurements in unexplored regions of the computational
domain, where they could significantly reduce uncertainty.

The issue with clusterization is evident in
Fig.~\ref{fig:clusterization_illustration}, where most clustered
designs shown place little to no measurement weight in the center of
the domain, where prior uncertainty is largest. It seems as though a
clustered design assumes that focusing measurement efforts on a small
subset of the domain can somehow yield information about the entire
domain. Intuitively, this is an unreasonable assumption, albeit it
holds for \emph{linear} models.

The view that clusterization is undesirable prompted the exploration
of various remedies to address this issue. One approach involves
merging close measurements \cite{fedorov1997}; however, this strategy
merely overlooks the phenomenon of measurement clusterization. An
alternative solution lies in \emph{clusterization-free design}s, where
measurement locations are deliberately chosen to be distant from one
another. This can be achieved by imposing distance constraints between
measurements or by introducing correlated errors that account for both
observation error and model misspecification \cite{Ucinski05}. For
instance, in the context of time-series analysis for pharmacokinetic
experiments, measurement clusterization can be mitigated by
incorporating the modeling of auto-correlation time within the noise
terms \cite{nyberg2012}.
